Submission and Formatting Instructions for the Twenty-Eighth \\ International Conference on Machine Learning (ICML 2012)

### 2012-04-10 klein2012/main.tex
Correction de typos repérées à la relecture du papier

% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage{icml2012}
\mathtoolsset{showonlyrefs=true}
\icmltitlerunning{Submission and Formatting Instructions for ICML 2012}

\newtheorem{definition}{Définition}
\newtheorem{lemme}{Lemme}
\newtheorem{theorem}{Théorème}
\newtheorem{hypo}{Hypothèse}
\newtheorem{cor}{Corollaire}
\newtheorem{prop}{Proposition}
\newtheorem{exemple}{Exemple}
\newtheorem{remarques}{Remarques}
\newtheorem{remarque}{Remarque}
\newtheorem{problem}{Problème}
\newtheorem{algo}{Algorithme}
\newtheorem{preuve}{Démonstration}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\argmin}{\operatorname*{argmin}}
\newcommand{\minp}{\operatorname*{min_+}}
\newcommand{\dom}{\operatorname*{dom}}
\newcommand{\Ker}{\operatorname*{Ker}}
\newcommand{\trace}{\operatorname*{trace}}
\newcommand{\cov}{\operatorname{cov}}
\newcommand{\epi}{\operatorname{epi}}
\newcommand{\card}{\operatorname*{Card}}
\newcommand{\conv}{\operatorname*{Conv}}
\newcommand{\vect}{\operatorname*{Vect}}
\newcommand{\var}{\operatorname{Var}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\erf}{\operatorname{erf}}
\newcommand{\bound}{\operatorname*{bound}}
\newcommand{\vpi}{\operatorname{VPI}}
\newcommand{\gn}{\operatorname{Gain}}
\newcommand{\p}{\operatorname{Pr}}
\newcommand{\mlp}{\operatorname{MLP}}
\newcommand*\tto[2]{\smash{\mathop{\longrightarrow}\limits_{#1}^{#2}}}
\newcommand*\ntto[2]{\smash{\mathop{\nrightarrow}\limits_{#1}^{#2}}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\ba}{\mathbf{a}}
\newcommand{\bb}{\mathbf{b}}
\newcommand{\bc}{\mathbf{c}}
\newcommand{\bd}{\mathbf{d}}
\newcommand{\be}{\mathbf{e}}
\newcommand{\af}{\mathbf{f}}
\newcommand{\bg}{\mathbf{g}}
\newcommand{\bh}{\mathbf{h}}
\newcommand{\bi}{\mathbf{i}}
\newcommand{\bj}{\mathbf{j}}
\newcommand{\bk}{\mathbf{k}}
\newcommand{\bl}{\mathbf{l}}
\newcommand{\bm}{\mathbf{m}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\bo}{\mathbf{o}}
\newcommand{\bp}{\mathbf{p}}
\newcommand{\bq}{\mathbf{q}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bs}{\mathbf{s}}
\newcommand{\bt}{\mathbf{t}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\bz}{\mathbf{z}}
\newcommand{\ma}{\mathbf{A}}
\newcommand{\mb}{\mathbf{B}}
\newcommand{\mc}{\mathbf{C}}
\newcommand{\md}{\mathbf{D}}
\newcommand{\me}{\mathbf{E}}
\newcommand{\mf}{\mathbf{F}}
\newcommand{\mg}{\mathbf{G}}
\newcommand{\mh}{\mathbf{H}}
\newcommand{\mi}{\mathbf{I}}
\newcommand{\mj}{\mathbf{J}}
\newcommand{\mk}{\mathbf{K}}
\newcommand{\ml}{\mathbf{L}}
\newcommand{\mm}{\mathbf{M}}
\newcommand{\mn}{\mathbf{N}}
\newcommand{\mo}{\mathbf{O}}
\newcommand{\Mp}{\mathbf{P}}
\newcommand{\mq}{\mathbf{Q}}
\newcommand{\mr}{\mathbf{R}}
\newcommand{\ms}{\mathbf{S}}
\newcommand{\mt}{\mathbf{T}}
\newcommand{\Mu}{\mathbf{U}}
\newcommand{\mv}{\mathbf{V}}
\newcommand{\mw}{\mathbf{W}}
\newcommand{\mx}{\mathbf{X}}
\newcommand{\my}{\mathbf{Y}}
\newcommand{\mz}{\mathbf{Z}}
\newcommand{\tphi}{\tilde{\Phi}}
\newcommand{\espace}{\text{ }}
\newcommand{\x}{\mathbf{x}}
\newcommand{\s}{\mathbf{s}}
\newcommand{\n}{\mathbf{n}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\rr}{\mathbf{r}}

@@

%\title{Classification structurée pour l'apprentissage par renforcement inverse}
%
%\author{\fontsize{12}{12}\selectfont{Edouard Klein\inst{1}$^,$\inst{2}, Bilal Piot \inst{1}$^,$\inst{3}, Matthieu Geist\inst{1}, Olivier Pietquin\inst{1}$^,$\inst{3}}}
%
%\institute{
%Sup\'elec, IMS Research group, France, \texttt{prenom.nom@supelec.fr}
%\and
%Equipe ABC,
%LORIA, France
%\and
%UMI 2958
%GeorgiaTech-CNRS, France
%}

@@

\twocolumn[
\icmltitle{Submission and Formatting Instructions for the Twenty-Eighth \\
           International Conference on Machine Learning (ICML 2012)}

@@

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boring formatting information, machine learning, ICML}

@@

\begin{abstract}
In this paper we tackle the problem of apprenticeship learning in the Inverse Reinforcement Learning (\emph{IRL}) framework.
An expert fulfils a task to be imitated by an artificial agent. The IRL framework assumes that the expert is acting optimally with respect to a reward function;
the goal is to guess this function from trajectories drawn by the expert. Existing IRL algorithms need one or more of the following items: complete trajectories from the expert,
a generative model of the environment for Monte-Carlo estimations, the knowledge of the transition probabilities, the ability to repeatedly solve the forward problem (the Reinforcement Learning (\emph{RL}) problem),
the access to the expert's strategy anywhere in the state space, or at the very least random samples covering a large part of it. Our contribution consists in a new IRL algorithm lifting all these constraints.
Using a supervised technique in which the structure of the underlying Markov Decision Process (\emph{MDP}) is implicitely injected, we give rise to a subgradient descent-based algorithm whose computational
and sample complexities are low. The sample base from the expert is the only data we need. A formal analysis of the algorithm as well as empirical demonstrations on two classic but challenging problems are provided.
\end{abstract}
%%%%%
\section{Introduction}
%%%%%
%Dans cette contribution, nous nous plaçons dans le contexte de l'apprentissage par démonstration. Dans ce cadre, un agent artificiel apprend à reproduire une tâche grâce à l'observation d'un expert réalisant cette tâche de manière optimale. Pour trouver une solution à ce problème, nous nous proposons d'utiliser le paradigme de l'apprentissage par renforcement inverse. On suppose alors que l'expert agit de manière à être récompensé pour son comportement relativement à une fonction de récompense inconnue. Le but de l'agent est alors d'inférer cette fonction des démonstrations données par l'expert pour ensuite optimiser son comportement relativement à cette fonction lui aussi.\\
In this paper, we consider the apprenticeship learning problem (also called learning by watching or learning from demonstrations problem) where an artificial agent (or apprentice) is learning to perform a task thanks to the observation of an expert demonstrating the task. In order to find a solution to this problem, we propose to use the inverse reinforcement learning paradigm where we suppose that the expert acts in order to be rewarded by an unknown reward function. The aim of the agent is then to infer the unknown reward function from the demonstrations given by the expert in order to find an optimal behavior (or policy) with respect to the reward function he has just inferred.
%La littérature sur le sujet est assez récente et trouve sa genèse dans \cite{russell1998learning}. Plusieurs directions ont depuis été explorées à partir de deux articles fondateurs \cite{ng2000algorithms} et \cite{abbeel2004apprenticeship}.\\
%Dans cette littérature (voir \cite{neu2009training} pour un bon survol), plusieurs verrous sont identifiés mais souvent ignorés pour proposer des algorithmes de résolution. Particulièrement, on suppose souvent connue la dynamique de l'environnement dans lequel évoluent l'expert et l'agent ou bien la mise à disposition de cet environnement (ou d'un simulateur) pour y tester les effets d'une politique quelconque. De plus, ces approches supposent de résoudre le problème direct (optimiser une politique pour une récompense donnée) un grand nombre de fois pour résoudre le problème inverse (trouver une récompense expliquant la politique). De plus, certaines de ces approches donnent en sortie une politique généralisant celle de l'expert plutôt qu'une récompense l'expliquant. Ces contraintes sont souvent incompatibles avec la mise en oeuvre des algorithmes proposés sur des cas réels.\\
The literature on the subject is quite recent and finds its origin in \cite{russell1998learning}. Different approaches have been explored since the first articles on the matter \cite{ng2000algorithms} and \cite{abbeel2004apprenticeship}.\\
In this literature (see \cite{neu2009training} for a survey), several difficulties are identified but often ignored in order to find algorithms. Specifically, we often suppose the dynamic of the environment where the expert and the agent are evolving known, or that we dispose of an environment model (via a simulator) in order to test the effects of a policy. Moreover, these approaches resolve the reinforcement learning problem (find an optimal policy knowing a given reward) an important number of times in order to resolve the inverse reinforcement problem (find a reward explaining the expert policy). Finally, some of these approaches give a policy that performs as well as the expert with respect to the unknown reward function but do not give a reward able to explain the expert policy. These constraints can be incompatible with the application of the proposed algorithms in real cases.
%Dans cette contribution, nous proposons une méthode de résolution du problème d'apprentissage par renforcement inverse s'affranchissant de toute connaissance additionnelle aux démonstrations fournies par l'expert. Nous proposons une preuve de convergence de cet algorithme et fournissons une expérimentation sur le problème du pendule inversé. Le choix de cette application se justifie par la difficulté  qu'elle pose pour les méthodes de résolution de la littérature.
In this paper, we propose a resolution method of the inverse reinforcement problem that is free from any other knowledge than the demonstrations given by the expert. We propose a proof of the convergence of our algorithm and give an experimentation on the inverted pendulum problem. The choice of this application is justified by the fact that it is difficult to resolve it by the other methods in the literature.
\section{Preliminaries}
%%%%%
\label{back.sec}
\subsection{Reinforcement Learning (RL) and Inverse Reinforcement Learning (IRL)}
%Le cadre dans lequel nous plaçons notre étude est celui de la prise de décisions séquentielles. La configuration du système à contrôler est alors complètement décrite à chaque instant discret $t$ par un état $s_t \in S$. Confronté à cet état, l'agent doit prendre un décision $a_t\in A$. Le système évolue alors vers l'état suivant $s_{t+1}$ selon une certaine probabilité de transition markovienne $p(s_{t+1}|s_t, a_t)$. Une politique de contrôle déterministe $\pi: S\rightarrow A$ définit le comportement d'un agent confronté à un tel problème de décisions séquentielles.\\
The general framework chosen is the sequential decision making problem posed in the Markov decision process (MDP) formalism. The configuration of the system is completely described at each discrete time $t$ by the state $s_t \in S$ where the agent must take a decision $a_t\in A$. Then, the system will evolve to the next state $s_{t+1} \in S$ with the probability $p(s_{t+1}|s_t, a_t)$. A deterministic policy which is a function $\pi: S\rightarrow A$ allows us to define the behavior of an agent faced to a sequential decision making problem. In this paper $S$ and $A$ will be finite sets and $\{P_{sa}\}_{s\in S,a\in A}$ will be a set of probability distributions, from $S$ to $[0,1]$ such that $P_{sa}(s')=p(s'|s, a)$, called transition probabilities.

@@

In the RL paradigm, the choice of a policy $\pi$ depends on a reward function $R: S \rightarrow \mathbb{R}$ which defines the immediate reward $R(s)$ given to the system when it attains the state $s \in S$. In this paradigm, the agent, following a policy $\pi$, is not going to learn to maximize the immediate reward but a criterion taking into account the futures rewards reached by the agent. That allows to specify the goal without specifying a way to reach it. The criterion commonly chosen is the value function $V^\pi$:

@@

\label{Vdef.eqn}
V^\pi(s) = E^\pi_s[\sum_{t\geq0}\gamma^tR(s_t)]=R(s) + \gamma\sum_{s'\in S}P_{s\pi(s)} V^\pi(s'),
where $\gamma \in [0,1)$ is a discount factor, and $E^\pi_s$ is the expectation over the distribution of the state sequences $(s_0,s_1,\dots)$ we pass through when we execute the policy $\pi$ starting from $s_0=s$. It is also possible to define the action-value function $Q^\pi$ which adds a degree of freedom in the choice of the first action:
\label{Qdef.eqn}
Q^\pi(s,a) = E^\pi_{sa}[\sum_{t\geq0}\gamma^tR(s_t)]=R(s) + \gamma\sum_{s'\in S}P_{sa} V^\pi(s'),
where $E^\pi_{sa}$ is the expectation over the distribution of the state sequences $(s_0,s_1,\dots)$ we pass through when we start from $s_0=s$, choose the action $a_0=a$ and then follow the policy $\pi$.
An optimal policy $\pi^*$ is defined as a policy which value function $V^{\pi*}=V^*=\sup_{\pi}V^{\pi}$ which means $\forall \pi, \forall s, V^{\pi*}(s) \geq V^\pi(s)$. An optimal policy $\pi^*$ can be computed easily thanks to the optimal action-value function $Q^{\pi*}=Q^*=\sup_{\pi}Q^\pi$ via a greedy mechanism:
\label{greedy.eqn}
\pi^*(s) \in \arg\max_{a\in A} Q^*(s,a).

@@

The tuple $M=\{S,A,\{P_{sa}\}_{s\in S,a\in A},\gamma,R\}$ containing the finite state space $S$, the finite action space $A$, the transitions probabilities $\{P_{sa}\}_{s\in S,a\in A}$, the discount factor $\gamma$ and the reward function $R$ is called a finite Markov decision process. Resolving the RL problem with respect to the criterion $V^\pi$  means to find $\pi^*$. In a finite MDP the existence of an optimal deterministic policy is guaranteed when the reward function $R$ is bounded.\\

@@

Sometimes, specifying manually the reward function can be very difficult while it is natural to an operator to control the system in order to complete the desired task. An exemple of this kind of task is driving a car. We can accomplish this task very easily in our daily-life but it would be very tricky to specify the precise weights given to each important driving-parameters such as the distance we should keep between the car in front of us, or maintaining a reasonable speed and so on. In such cases, it is useful to infer the reward function thanks to observed behavior.\\

@@

That is the aim of the Inverse Reinforcement Learning. The problem is to find the reward function optimized by an expert. Usually, the expert is considered as an optimal agent in a MDP and we have access to the expert optimal policy $\pi_E$ or to some trajectories of this policy.

@@

This problem is ill-posed in the sense that there is not uniqueness of the reward function for which the expert policy is optimal (\cite{ng1999policy}). Particularly, each policy is optimal with respect to the uniformly-zero reward function. One should favor some solutions.

@@

The reward function $R$ is the unknown of the problem. We can suppose that it is possible to dispose of $p$ features via $\psi: S\rightarrow \mathbb{R}^p$, which is a functions from $S$ to $\mathbb{R}^p$ and each of the components of $\psi$: $\{\psi_i\}_{1\leq i \leq p}$ is one feature, such that:

@@

\label{hatRdef.eqn}
\exists \theta | R(s) = \theta^T\psi(s) = \sum_{i=1}^p\theta_i\psi_i(s).

@@

This constraint can be relaxed if necessary. Putting this expression of the reward function in the definition of the action-value function (Eq. \eqref{Qdef.eqn}) allows us to define the term $\mu^\pi$ which is often used in the existing algorithms:

@@

Q^\pi(s,a)=\theta^T\mu^\pi(s,a),\text{ }\mu^\pi(s,a) = E^\pi_{sa}[\sum_t\gamma^t\psi(s_t)].
\label{Qmu.eqn}

@@

The term $\mu^\pi$ is called the feature expectation of the policy $\pi$. We can observe that the chosen parametrization of the reward function $R$ and the dynamic resulting from the application of the policy $\pi$ in the MDP impose the parametrization of $Q^\pi$. To shorten notations, we may use $\mu^\pi_{sa}$ as $\mu^\pi(s,a)$.\\

>>>>

% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{url}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage{icml2012}
\mathtoolsset{showonlyrefs=true}

\newtheorem{definition}{Définition}
\newtheorem{lemme}{Lemma}
\newtheorem{theorem}{Théorème}
\newtheorem{hypo}{Hypothèse}
\newtheorem{cor}{Corollaire}
\newtheorem{prop}{Proposition}
\newtheorem{exemple}{Exemple}
\newtheorem{remarques}{Remarques}
\newtheorem{remarque}{Remarque}
\newtheorem{assumption}{Assumption}
\newtheorem{problem}{Problème}
\newtheorem{algo}{Algorithme}
\newtheorem{preuve}{Démonstration}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\argmin}{\operatorname*{argmin}}
\newcommand{\minp}{\operatorname*{min_+}}
\newcommand{\dom}{\operatorname*{dom}}
\newcommand{\Ker}{\operatorname*{Ker}}
\newcommand{\trace}{\operatorname*{trace}}
\newcommand{\cov}{\operatorname{cov}}
\newcommand{\epi}{\operatorname{epi}}
\newcommand{\card}{\operatorname*{Card}}
\newcommand{\conv}{\operatorname*{Conv}}
\newcommand{\vect}{\operatorname*{Vect}}
\newcommand{\var}{\operatorname{Var}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\erf}{\operatorname{erf}}
\newcommand{\bound}{\operatorname*{bound}}
\newcommand{\vpi}{\operatorname{VPI}}
\newcommand{\gn}{\operatorname{Gain}}
\newcommand{\p}{\operatorname{Pr}}
\newcommand{\mlp}{\operatorname{MLP}}
\newcommand*\tto[2]{\smash{\mathop{\longrightarrow}\limits_{#1}^{#2}}}
\newcommand*\ntto[2]{\smash{\mathop{\nrightarrow}\limits_{#1}^{#2}}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\ba}{\mathbf{a}}
\newcommand{\bb}{\mathbf{b}}
\newcommand{\bc}{\mathbf{c}}
\newcommand{\bd}{\mathbf{d}}
\newcommand{\be}{\mathbf{e}}
\newcommand{\af}{\mathbf{f}}
\newcommand{\bg}{\mathbf{g}}
\newcommand{\bh}{\mathbf{h}}
\newcommand{\bi}{\mathbf{i}}
\newcommand{\bj}{\mathbf{j}}
\newcommand{\bk}{\mathbf{k}}
\newcommand{\bl}{\mathbf{l}}
\newcommand{\bm}{\mathbf{m}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\bo}{\mathbf{o}}
\newcommand{\bp}{\mathbf{p}}
\newcommand{\bq}{\mathbf{q}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bs}{\mathbf{s}}
\newcommand{\bt}{\mathbf{t}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\bz}{\mathbf{z}}
\newcommand{\ma}{\mathbf{A}}
\newcommand{\mb}{\mathbf{B}}
\newcommand{\mc}{\mathbf{C}}
\newcommand{\md}{\mathbf{D}}
\newcommand{\me}{\mathbf{E}}
\newcommand{\mf}{\mathbf{F}}
\newcommand{\mg}{\mathbf{G}}
\newcommand{\mh}{\mathbf{H}}
\newcommand{\mi}{\mathbf{I}}
\newcommand{\mj}{\mathbf{J}}
\newcommand{\mk}{\mathbf{K}}
\newcommand{\ml}{\mathbf{L}}
\newcommand{\mm}{\mathbf{M}}
\newcommand{\mn}{\mathbf{N}}
\newcommand{\mo}{\mathbf{O}}
\newcommand{\Mp}{\mathbf{P}}
\newcommand{\mq}{\mathbf{Q}}
\newcommand{\mr}{\mathbf{R}}
\newcommand{\ms}{\mathbf{S}}
\newcommand{\mt}{\mathbf{T}}
\newcommand{\Mu}{\mathbf{U}}
\newcommand{\mv}{\mathbf{V}}
\newcommand{\mw}{\mathbf{W}}
\newcommand{\mx}{\mathbf{X}}
\newcommand{\my}{\mathbf{Y}}
\newcommand{\mz}{\mathbf{Z}}
\newcommand{\tphi}{\tilde{\Phi}}
\newcommand{\espace}{\text{ }}

\newcommand{\x}{\mathbf{x}}
\newcommand{\s}{\mathbf{s}}
\newcommand{\n}{\mathbf{n}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\rr}{\mathbf{r}}

@@

%\title{Classification structurée pour l'apprentissage par renforcement inverse}
%
%\author{\fontsize{12}{12}\selectfont{Edouard Klein\inst{1}$^,$\inst{2}, Bilal Piot \inst{1}$^,$\inst{3}, Matthieu Geist\inst{1}, Olivier Pietquin\inst{1}$^,$\inst{3}}}
%
%\institute{
%Sup\'elec, IMS Research group, France, \texttt{prenom.nom@supelec.fr}
%\and
%Equipe ABC,
%LORIA, France
%\and
%UMI 2958
%GeorgiaTech-CNRS, France
%}
\icmltitlerunning{Structured Classification for Inverse
Reinforcement Learning}

@@

\twocolumn[ \icmltitle{Structured Classification for Inverse
Reinforcement Learning}

@@

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{inverse reinforcement learning, structured multi-class
classification}

@@

\begin{abstract}
This paper addresses the Inverse Reinforcement Learning (IRL)
problem which is a particular case of learning from demonstrations.
The IRL framework assumes that an expert, demonstrating a task, is
acting optimally with respect to an unknown reward function to be
discovered. Unlike most of existing IRL algorithms, the proposed
approach doesn't require any of the following: complete trajectories
from the expert, a generative model of the environment, the
knowledge of the transition probabilities, the ability to repeatedly
solve the forward Reinforcement Learning (RL) problem, the expert's
policy anywhere in the state space. Using a classification approach
in which the structure of the underlying Markov Decision Process
(MDP) is implicitly injected, we end-up with an efficient
subgradient descent-based algorithm.
% whose computational and sample
%complexities are low.
In addition, only a small amount of expert
demonstrations (not even in the form of trajectories but simple
transitions) is required.
%A formal analysis of the algorithm as well as empirical
%demonstrations on two classic but challenging problems are provided.
\end{abstract}
%%%%%
\section{Introduction}
%%%%%
%Dans cette contribution, nous nous plaçons dans le contexte de l'apprentissage par démonstration. Dans ce cadre, un agent artificiel apprend à reproduire une tâche grâce à l'observation d'un expert réalisant cette tâche de manière optimale. Pour trouver une solution à ce problème, nous nous proposons d'utiliser le paradigme de l'apprentissage par renforcement inverse. On suppose alors que l'expert agit de manière à être récompensé pour son comportement relativement à une fonction de récompense inconnue. Le but de l'agent est alors d'inférer cette fonction des démonstrations données par l'expert pour ensuite optimiser son comportement relativement à cette fonction lui aussi.\\

This contribution addresses the problem of learning by watching or
learning from demonstrations. In this context, an artificial agent
learns to perform a task by observing an expert demonstrating the
task. In order to find a solution to this problem, the Inverse
Reinforcement Learning (IRL) paradigm has been
proposed~\cite{russell1998learning}. The fundamental idea of IRL is
that the expert acts so as to be rewarded by an unknown reward
function. The aim of the IRL agent is then to infer the unknown
reward function from the demonstrations given by the expert, the
reward being considered as the most compact representation of the
task to be transferred to the agent. This way, an optimal behavior
can be obtained in any situation by optimizing the action selection
process according to the reward function. Unlike supervised
learning, the agent may not purely imitate the expert's actions but
will try to accomplish the same task optimally.
%
%
%La littérature sur le sujet est assez récente et trouve sa genèse dans \cite{russell1998learning}. Plusieurs directions ont depuis été explorées à partir de deux articles fondateurs \cite{ng2000algorithms} et \cite{abbeel2004apprenticeship}.\\
%Dans cette littérature (voir \cite{neu2009training} pour un bon survol), plusieurs verrous sont identifiés mais souvent ignorés pour proposer des algorithmes de résolution. Particulièrement, on suppose souvent connue la dynamique de l'environnement dans lequel évoluent l'expert et l'agent ou bien la mise à disposition de cet environnement (ou d'un simulateur) pour y tester les effets d'une politique quelconque. De plus, ces approches supposent de résoudre le problème direct (optimiser une politique pour une récompense donnée) un grand nombre de fois pour résoudre le problème inverse (trouver une récompense expliquant la politique). De plus, certaines de ces approches donnent en sortie une politique généralisant celle de l'expert plutôt qu'une récompense l'expliquant. Ces contraintes sont souvent incompatibles avec la mise en oeuvre des algorithmes proposés sur des cas réels.\\
%
%
The literature on the subject is quite recent and finds its origin
in \cite{russell1998learning}. Different approaches have been
explored since the first articles on the matter
\cite{ng2000algorithms,abbeel2004apprenticeship}.

In this literature (see \cite{neu2009training} for a survey),
several bottlenecks are identified but often ignored when developing
algorithms. Specifically, the dynamic of the environment where the
expert and the agent are evolving is assumed to be known, or an
environment model (via a simulator) must be available so as to test
the effects of a policy. Moreover, most of existing approaches
require solving the direct reinforcement learning problem (find an
optimal policy knowing a given reward) several times. Finally, in
many cases, the output of the algorithm is a policy but not a unique
reward function which makes task transfer hard to realize. These
constraints can lead to the development of algorithms unapplicable
to real problems.
%Dans cette contribution, nous proposons une méthode de résolution du problème d'apprentissage par renforcement inverse s'affranchissant de toute connaissance additionnelle aux démonstrations fournies par l'expert. Nous proposons une preuve de convergence de cet algorithme et fournissons une expérimentation sur le problème du pendule inversé. Le choix de cette application se justifie par la difficulté  qu'elle pose pour les méthodes de résolution de la littérature.
%
In this paper, we propose a resolution method of the inverse
reinforcement problem that is free from any other knowledge than the
demonstrations given by the expert. The proposed approach, the
Structured Classification for IRL (SCIRL), relies on a
classification paradigm in which the structure of the environment's
dynamics is implicitly introduced. Sec.~\ref{back.sec}
and~\ref{biblio.sec} provide the necessary background and review a
part of the state of the art. Sec.~\ref{sec:scirl} introduces our
contribution and Sec.~\ref{proof.sec} analyses the asymptotic
algorithm. Section~\ref{exp.sec} shows results obtained on two
classical benchmarks: inverted pendulum and the highway driving
problems.
%
%A proof of the convergence of the algorithm is provided in
%Section~\ref{proof.sec} and Section~\ref{exp.sec} shows results
%obtained on two classical benchmarks: inverted pendulum and the
%highway driving problems. The choice of the first application is
%justified by the fact that it is hard to solve by the other methods.

%%%%
\section{Background}
%%%%%
\label{back.sec}
%%%
Let's start with some background on RL and IRL and settle the
notations.

\subsection{(Inverse) Reinforcement Learning}
%%%
%Le cadre dans lequel nous plaçons notre étude est celui de la prise de décisions séquentielles. La configuration du système à contrôler est alors complètement décrite à chaque instant discret $t$ par un état $s_t \in S$. Confronté à cet état, l'agent doit prendre un décision $a_t\in A$. Le système évolue alors vers l'état suivant $s_{t+1}$ selon une certaine probabilité de transition markovienne $p(s_{t+1}|s_t, a_t)$. Une politique de contrôle déterministe $\pi: S\rightarrow A$ définit le comportement d'un agent confronté à un tel problème de décisions séquentielles.\\

The problem is seen as a sequential decision making problem posed in
the Markov Decision Process (MDP) formalism. In this formalism, a
system to be controlled can reach different configurations
completely described at each discrete time $t$ by a state $s_t \in
S$. Given the knowledge of the state, the agent must take a decision
$a_t\in A$. Then, the system evolves to a next state $s_{t+1} \in S$
with the Markovian probability $p(s_{t+1}|s_t, a_t)$. %(in short
%$p(s'|s,a)$ where $s'=s_{t+1}, s=s_t$ and $a=a_t$).
A policy, which
is a function $\pi: S\rightarrow A$, defines the behavior of the
agent. In this paper all the policies considered are deterministic,
$S$ and $A$ will be finite sets and $\{P_{sa}\}_{s\in S,a\in A}$
will be a set of probability distributions, from $S$ to $[0,1]$ such
that $P_{sa}(s')=p(s'|s, a)$, called transition probabilities.

@@

The RL paradigm aims at solving the problem of finding the optimal
sequence of actions in an MDP. The optimality criterion is based on
the definition of a reward function $R: S \rightarrow \mathbb{R}$
which defines the immediate reward $R(s)$ given to the agent when
the system is in the state $s \in S$. The agent should not learn to
maximize the immediate reward but a criterion taking into account
future rewards. The criterion commonly chosen is the value function
$V^\pi$:
\begin{align}
\label{Vdef.eqn} V^\pi(s) & = E^\pi_s[\sum_{t\geq0}\gamma^tR(s_t)]\\
& = R(s) + \gamma\sum_{s'\in S}P_{s\pi(s)}(s') V^\pi(s'),
\end{align}
where $\gamma \in [0,1)$ is a discount factor, and $E^\pi_s$ is the
expectation over the distribution of the state sequences
$(s_0,s_1,\dots)$ the agent passes through when executing the policy
$\pi$ starting from $s_0=s$. One can define the action-value
function $Q^\pi$ which adds a degree of freedom in the choice of the
first action:
\begin{align}
\label{Qdef.eqn} Q^\pi(s,a) & =
E^\pi_{sa}[\sum_{t\geq0}\gamma^tR(s_t)]\\
& = R(s) + \gamma\sum_{s'\in S}P_{sa}(s') V^\pi(s').
\end{align}
%where $E^\pi_{sa}$ is the expectation over the distribution of the
%state sequences $(s_0,s_1,\dots)$ we pass through when we start from
%$s_0=s$, choose the action $a_0=a$ and then follow the policy $\pi$.
An optimal policy $\pi^*$ is defined as a policy with maximal value
function, $V^{\pi*}=V^*=\sup_{\pi}V^{\pi}$, which means that
$\forall \pi, \forall s, V^{\pi*}(s) \geq V^\pi(s)$. An optimal
policy $\pi^*$ can be computed easily thanks to the optimal
action-value function $Q^{\pi*}=Q^*=\sup_{\pi}Q^\pi$ via a greedy
mechanism:

@@

\label{greedy.eqn} \pi^*(s) \in \arg\max_{a\in A} Q^*(s,a).

@@

The tuple $M=\{S,A,\{P_{sa}\}_{s\in S,a\in A},\gamma,R\}$ is called
a finite MDP. Resolving the RL problem with respect to the criterion
$V^\pi$  means to find $\pi^*$, usually without the full knowledge
of the MDP but only by observing transitions and rewards. In a
finite MDP the existence of an optimal deterministic policy is
guaranteed when the reward function $R$ is bounded.

@@

The optimal behaviour being defined relatively to the reward, its
observation is mandatory to learn. Sometimes, specifying manually
the reward function can be a hard job while it is natural to an
operator to achieve the desired task. For instance, driving a car is
such a task. One can accomplish this task very easily in his
daily-life but it is very tricky to specify the precise weights
given to each important driving-parameters such as the distance to
keep with the leading car, the speed below which it is unreasonable
to drive and so on. In such cases, inferring the reward function
from observed behavior would be helpful.

@@

It is the aim of IRL. The expert, demonstrating the task, is thus
considered as an optimal agent in a MDP and the IRL agent has access
to the expert optimal policy $\pi_E$ or to some trajectories drawn
from this policy.

@@

The problem of finding a reward function from expert demonstrations
in the IRL framework is ill-posed. Indeed, there is not uniqueness
of the reward function for which the expert policy is
optimal~\cite{ng1999policy}. Particularly, each policy is
optimal with respect to the uniformly-zero reward function. Some
solutions have to be favored.

@@

The actual expert reward function $R_E$ is unknown. The problem is
 thus to find a reward function $R_\theta$ for which the policy of
the expert is also optimal. Let's assume that, given a set of $p$
features $\psi: S\rightarrow \mathbb{R}^p$, each components of
$\psi$: $\{\psi_i\}_{1\leq i \leq p}$ being a basis function from
$S$ to $\mathbb{R}$, such a reward can be written:

@@

\label{hatRdef.eqn} %\exists \theta |
R_\theta(s) = \theta^T\psi(s) =
\sum_{i=1}^p\theta_i\psi_i(s).

@@

Using this expression allows rewriting (Eq. \eqref{Qdef.eqn}) as:

@@

Q^\pi(s,a)=\theta^T\mu^\pi(s,a),\text{ }\mu^\pi(s,a) =
E^\pi_{sa}[\sum_{t\geq 0}\gamma^t\psi(s_t)]. \label{Qmu.eqn}

@@

The term $\mu^\pi$ is called the feature expectation of the policy
$\pi$. We can observe that the chosen parametrization of the reward
function $R_\theta$ and the dynamics resulting from the application
of the policy $\pi$ in the MDP gives rise to an induced
parametrization for $Q^\pi$. To
